Trawl Summary

### 2020-05-11 text/trawl_summary.Rmd
updated trawl_summary with species sampling summaries.

---
title: 'Trawl Summary'
author: "Jim Junker"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
      keep_md: true
      toc: true
      toc_float: true
      toc_depth: 3
      numbered_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vegan)
library(betapart)
library(lubridate)

@@

These data sets range in coverage, and are being added to continuously:

@@

>>>>

---
title: 'Trawl Summary'
author: "Jim Junker"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
      out_dir <- '../docs';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir, 'trawl_summary.html')) })
output: 
  html_document:
      keep_md: true
      toc: true
      toc_float: true
      toc_depth: 3
      numbered_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vegan)
library(betapart)
library(lubridate)
library(viridis)

'%ni%' <- Negate('%in%')

@@

These data vary in coverage and are being updated:

@@

There are a lot of sampling trawls over the past decade (n = 359) and they are spread pretty consistently over the year, but with some periods of heavy sampling:

```{r month sampling events, include=TRUE, echo=FALSE}

TB_month_hist <- TB_trawl_samplings %>%
  mutate(month_num = lubridate::month(Date), 
         month_name = lubridate::month(Date, label = TRUE))

  ggplot(TB_month_hist, aes(month_num)) + geom_bar(width = 1) +
    scale_x_continuous(breaks = seq(1,12, by = 1), labels = lubridate::month(seq(1,12,by = 1), label = TRUE)) +
    theme_minimal() +
    theme(axis.title = element_blank())

## Species sampling

Over the full dataset we can estimate the species accumulation by # of trawls

```{r full richness curve, include = TRUE, echo=FALSE}

# Full species accumulation curve (SAC)

##create taxa by site matrix
TB_trawl_commonsite <- TB_trawl_data %>%
  select(year, month, date_id, Common_name) %>%
  group_by(date_id, Common_name) %>%
  distinct() %>% 
  mutate(pres = 1) %>%
  pivot_wider(names_from = Common_name, values_from = pres, values_fill = list(pres = 0)) %>%
  ungroup()

sp1 <- vegan::specaccum(TB_trawl_commonsite %>% select(-year, -month, -date_id))
spp_mod.df <- data.frame(trawls = sp1[["sites"]], richness = sp1[["richness"]], richness_sd = sp1[["sd"]])

ggplot(spp_mod.df, aes(x = trawls, y = richness)) + geom_path(size = 2, colour = 'red') +
  geom_ribbon(aes(x = trawls, ymin = (richness - 1.96*richness_sd), ymax = (richness + 1.96*richness_sd)), colour = "red", fill = "red", alpha = 0.5) +
  scale_y_continuous(name = expression(Sigma~"species"), breaks = seq(0,roundUP(max(spp_mod.df$richness), to = 10), by = 10), labels = seq(0,roundUP(max(spp_mod.df$richness), to = 10), by = 10)) +
  scale_x_continuous(name = "Trawls", breaks = seq(0,roundUP(max(spp_mod.df$trawls), to = 10), by = 20)) +
  theme_minimal()

And by the trawls within each year to get at the differences in sampling effort (assuming each trawl is the same). This suggests there are certain years (or trawl locations):

```{r, include = TRUE, echo = FALSE, warning=FALSE}
## SAC for each year
TB_trawl_commondate <- TB_trawl_data %>%
  select(year, Date, Common_name) %>%
  group_by(Date, Common_name) %>%
  distinct() %>% 
  mutate(pres = 1) %>%
  pivot_wider(names_from = Common_name, values_from = pres, values_fill = list(pres = 0)) %>%
  ungroup()

TB_trawl_taxasite_yr <- TB_trawl_commondate %>%
  group_split(year) %>%
  map(., ~.x %>%  select(-year, -Date) %>% vegan::specaccum(.) %>%
       rlist::list.remove(names(.) %ni% c("sites","richness","sd")) %>%
        bind_cols) %>% 
  map2(., TB_trawl_commondate %>% group_split(year), ~..1 %>% 
         mutate(year = ..2 %>% select(year) %>% unlist %>% as.factor)) %>%
  bind_rows

ggplot(TB_trawl_taxasite_yr, aes(x = sites, y = richness, group = year, colour = year)) +
  geom_path(size = 1.5) +# geom_ribbon(aes(x = sites, ymin = (richness - 1.96*sd), ymax = (richness + 1.96*sd), fill = year), alpha = 0.2) +
  scale_y_continuous(name = expression(Sigma~"species"),breaks = seq(0, roundUP(max(TB_trawl_taxasite_yr$richness), to = 10), by = 10), labels = seq(0, roundUP(max(TB_trawl_taxasite_yr$richness), to = 10), by = 10)) +
  scale_x_continuous(name = "Trawls", breaks = seq(0, roundUP(max(TB_trawl_taxasite_yr$sites), to = 10), by = 5)) +
  scale_colour_viridis(discrete = TRUE) + 
  scale_fill_viridis(discrete = TRUE) +
  theme_minimal()

Squinting this suggests that, generally, there might be increasing # of species more broadly.
